---
title: "P8451 Assignment 1"
output: html_document
date: "2023-01-24"
---

```{r setup, include = FALSE}
library(tidyverse)
library(readr)

knitr::opts_chunk$set(
  warning = FALSE, 
  message = FALSE,
  fig.dim = c(12, 7))

theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

## 1. Descriptive statistics

### Reading in the data
First, let's read in the dataset.

```{r}
bc_data = read_csv("./bcdata_Assignment1.csv")
```

We can see that the `bc_data` table contains `r nrow(bc_data)` observations of `r ncol(bc_data)` variables.

### Summary table
We will construct a table providing summaries of the mean, median, minimum value, and maximum value of each variable in the dataset.

```{r}
summary(bc_data) %>% knitr::kable()
```

## 2. Recoding BMI
The code chunk below creates a new variable, `BMI_cat`, which recodes BMI into 7 categories as defined by the WHO.

```{r}
bc_data = bc_data %>% 
  mutate(BMI_cat = as.factor(case_when(BMI < 16.5 ~ "Severely underweight", 
                             BMI < 18.5 & BMI >= 16.5 ~ "Underweight", 
                             BMI < 25 & BMI >= 18.5 ~ "Normal weight", 
                             BMI < 30 & BMI >= 25 ~ "Overweight",
                             BMI < 35 & BMI >= 30 ~ "Obesity class I",
                             BMI < 40 & BMI >= 35 ~ "Obesity class II",
                             BMI >= 40 ~ "Obesity class III"
                             )))
```

## 3. Plotting breast cancer rates across BMI categories
Next, we will create a bar chart showing the proportion of breast cancer cases and controls within each BMI category. For interpretability, we will format the `Classification` variable into a factor variable such that 1 = Control and 2 = Case.

```{r plot}
bc_data = bc_data %>% 
  mutate(BMI_cat = fct_reorder(BMI_cat, BMI), 
         Classification = as.factor(case_when(Classification == 1 ~ "Control",
                                              Classification == 2 ~ "Case"))) %>% 
  mutate(Classification = fct_relevel(Classification, "Control", "Case"))

BC_BMI = bc_data %>% 
  group_by(BMI_cat, Classification) %>% 
  summarise(N = n()) %>% 
  mutate(Proportion = N/sum(N)) 

BC_BMI %>% ggplot(aes(x = BMI_cat, y = Proportion, fill = Classification)) + 
  geom_bar(position = position_stack(reverse = F), stat = "identity") +
  labs(
    x = "BMI Category",
    y = "Proportion",
    title = "Proportion of breast cancer cases and controls by BMI category") +
  scale_fill_brewer(palette = "PiG")
```

## 4. Logistic regression model of breast cancer
The code below constructs a logistic regression model using breast cancer classification as the outcome and glucose, HOMA-IR, leptin, BMI (continuous) and age as the independent variables.

```{r logistic}
fit_logistic = bc_data %>% 
  glm(Classification ~ Glucose + HOMA + Leptin + BMI + Age, data = ., family = binomial())

fit_logistic %>% 
  broom::tidy() %>% 
  filter(term == "HOMA") %>% 
  mutate(CIL = estimate - 1.96*std.error,
         CIU = estimate + 1.96*std.error) %>%
  select(term, estimate, CIL, CIU) %>% 
  rename("Term" = term,
         "Beta estimate" = estimate) %>% 
  knitr::kable(digits = 3)
```

We can see that for a 1-unit increase in HOMA-IR, the log odds of having breast cancer is expected to increase by 0.274 (95% CI: -0.063, 0.611). 

## 5. Linear regression model of breast cancer
Next, we will construct a linear regression model using insulin as the outcome and BMI (continuous), age, and glucose as the independent variables.

```{r linear}
fit_linear = bc_data %>% 
  lm(Insulin ~ BMI + Age + Glucose, data = .)

fit_linear %>% 
  broom::tidy() %>% 
  filter(term == "Age") %>% 
  mutate(CIL = estimate - 1.96*std.error,
         CIU = estimate + 1.96*std.error) %>%
  select(term, estimate, CIL, CIU) %>% 
  rename("Term" = term,
         "Beta estimate" = estimate) %>% 
  knitr::kable(digits = 3)
```

We can see that for a 1-year increase in age, the log odds of insulin increasing by 1 unit is expected to decrease by 0.054 (95% CI: -0.156, 0.048). 
